### 2022-04-23 chapters/chapter4.tex
update

\chapter{Diversity-based Trajectory and Goal Selection with Hindsight Experience Replay}
\section{Introduction}
Deep reinforcement learning (DRL)~\cite{arulkumaran2017deep}, in which neural networks are used as function approximators for reinforcement learning (RL), has been shown to be capable of solving complex control problems in several environments, including board games~\cite{schrittwieser2020mastering,silver2017mastering}, video games~\cite{berner2019dota,mnih2015human,vinyals2019grandmaster}, simulated and real robotic manipulation~\cite{andrychowicz2020learning,gu2017deep,levine2016end} and simulated autonomous driving~\cite{kiran2021deep}.

However, learning from a \textit{sparse} reward signal, where the only reward is provided upon the completion of a task, still remains difficult. An agent may rarely or never encounter positive examples from which to learn in a sparse-reward environment. Many domains therefore provide dense reward signals~\cite{brockman2016openai}, or practitioners may turn to reward shaping~\cite{ng1999theory}. Designing dense reward functions typically requires prior domain knowledge, making this approach difficult to generalise across different environments.

Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}. Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier. For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.

Hindsight experience replay (HER)~\cite{NIPS2017_453fadbd} was proposed to improve the learning efficiency of goal-oriented RL agents in sparse reward settings: when past experience is replayed to train the agent, the desired goal is replaced (in ``hindsight'') with the achieved goal, generating many positive experiences. In the above example, the desired target position would be overwritten with the achieved target position, with the achieved reward also being overwritten correspondingly.

We note that HER, whilst it enabled solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training. In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}. HER with energy-based prioritisation (HEBP)~\cite{zhao2018energy} assumes semantic knowledge about the goal-space and uses the energy of the target objects to sample trajectories with high energies, and then samples transitions uniformly. Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time. In this work, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Fig.~\ref{fig:illustration}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples. In this paper, DTGSH is evaluated on five challenging robotic manipulation tasks. From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
    \caption{Overview of DTGSH. Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer. During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
    \label{fig:illustration}
\end{figure}

\section{Methodology}
We now formally describe the two main components of our method, DTGSH: 1) a diversity-based trajectory selection module to sample valuable trajectories for the further goal selection; 2) a diversity-based goal selection module to select transitions with diverse goal states from the previously selected trajectories. Together, these select informative transitions from a large area of the goal space, improving the agent's ability to learn and generalise.

\subsection{Diversity-based Trajectory Selection}
We propose a diversity-based prioritization method to select valuable trajectories for efficient training. Related to HEBP's prioritisation of high-energy trajectories~\cite{zhao2018energy}, we hypothesise that trajectories that achieve diverse goal states $g^{ac}_{t}$ are more valuable for training; however, unlike HEBP, we do not require knowledge of the goal space semantics.

In a robotic manipulation task, the agent needs to move a target object from its initial position, $g^{ac}_{1}$, to the target position, $g$. If the agent never moves the object, despite hindsight relabelling it will not be learning information that would directly help in task completion. On the other hand, if the object moves a lot, hindsight relabelling will help the agent learn about meaningful interactions.

In our approach, DPPs are used to model the diversity of achieved goal states $g^{ac}_{t}$ in an episode, or subsets thereof. For a single trajectory $\Tau$ of length $T$, we divide it into several partial trajectories $\tau_{j}$ of length $b$, with achieved goal states $\{g^{ac}_{t}\}_{t=n:n+b-1}$. That is, with a sliding window of $b = 2$, a trajectory $\Tau$ can be divided into $N_p$ partial trajectories:
\begin{equation}
    \Tau_{i} = \{\{\underbrace{g^{ac}_{1}, g^{ac}_{2}}_{\tau_{1}}\}, \{\underbrace{g^{ac}_{2}, g^{ac}_{3}}_{\tau_{2}}\}, \{\underbrace{g^{ac}_{3}, g^{ac}_{4}}_{\tau_{3}}\}, \cdots, \{\underbrace{g^{ac}_{T-1}, g^{ac}_{T}}_{\tau_{N_{p}}}\}\}.
\end{equation}
The diversity $d_{\tau_{j}}$ of each partial trajectory $\tau_{j}$ can be computed as:
\begin{equation}
    d_{\tau_{j}} = \text{det}(L_{\tau_{j}}), 
\label{eq:partial_div}
\end{equation}
where $L_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
\begin{equation}
    L_{\tau_{j}} = M^{T}M,
\end{equation}
and $M=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}. Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
\begin{equation}
    d_\Tau = \sum_{j=1}^{N_{p}} d_{\tau_{j}}.
\label{eq:total_div}
\end{equation}

@@

\subsection{Diversity-based Goal Selection}
In prior work~\cite{NIPS2017_453fadbd,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training. However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information. In order to form a minibatch with diverse goals for more efficient learning, we use $k$-DPPs~\cite{kulesza2011k} for sampling goals. Compared to the standard DPP, a $k$-DPP is a conditional DPP where the subset $Y$ has a fixed size $k$, with the probability distribution function:
\begin{equation}
    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(L_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(L_{Y^{\prime}})}.
\end{equation}
$k$-DPPs are more appropriate for goal selection with a minibatch of fixed size $k$. Given $m > k$ trajectories sampled from the replay buffer, we first uniformly sample a transition from each of the $m$ trajectories. Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals''). Fig.~\ref{fig:goal_sampling} gives an example of uniform vs. $k$-DPP sampling, demonstrating the increased coverage of the latter. Fig.~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/goals_distribution_2d.pdf}
        \caption{Plot of candidate goals and selected goals. $k$-DPP sampling is more likely to sample points from the full span of the goal space.}
        \label{fig:goal_sampling}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/kde.pdf}
        \caption{Kernel density estimation of the distributions of goals. $k$-DPP leads to a more uniform selection of goals over the support of the goal space.}
        \label{fig:kde}
    \end{subfigure}
    \caption{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively. The candidate goals are distributed over a 2D ($xy$) space. Note that  $k$-DPP sampling (right hand plots) results in a broader span of selected goals in $xy$ space compared to uniform sampling (left hand plots).}
\end{figure}

% pseudo code of dpp-sampling
\begin{algorithm}[t!]
\caption{Diversity-based Goal Selection using $k$-DPP}
\label{alg:goal_sample}
\begin{algorithmic}[1]
 \REQUIRE set of $m$ candidate goal states $\mathcal{G}:=\{g_{i}\}_{i=1:m}$, minibatch size $k$
 \STATE $J \leftarrow\varnothing$, $M \leftarrow  [g_{1}, g_{2}, \cdots, g_{m}]$
 \STATE Calculate the DPP kernel matrix $L_M$
 \STATE $\{\boldsymbol{v}_{n}, \lambda_{n}\} \leftarrow \text{EigenDecomposition}(L_M)$
 %\STATE $e_{k}(\lambda_{1}, \cdots, \lambda_{m}): = \sum\limits_{|J'|=k}\prod_{n\in J'}\lambda_{n}$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \STATE $e_k(\lambda_1,\lambda_2,\dots,\lambda_m) := 
  \sum\limits_{J^{\prime} \subseteq \{1,2,\dots,m\}\atop|J^{\prime}|=k} \prod\limits_{n \in J^{\prime}} \lambda_n$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \FOR{$\mathrm{n} = m, m-1, \cdots, 1$}
 \IF{$u\sim \text{Uniform}[0, 1] < \lambda_{n}\frac{e^{n-1}_{k-1}}{e_{k}^{n}}$}
 \STATE $J \leftarrow J \cup \{n\}$, $k \leftarrow k - 1$
 \IF{$k = 0$}
 \STATE \textbf{break}
 \ENDIF
 \ENDIF
 \ENDFOR
 \STATE $V \leftarrow \{\boldsymbol{v}_{n}\}_{n \in J}$, $B \leftarrow \varnothing$
 \WHILE{$|V| > 0$}
 \STATE Select $g_{i}$ from $\mathcal{G}$ with $p(g_{i}):=\frac{1}{|V|}\sum_{\boldsymbol{v}\in V}(\boldsymbol{v}^{T}\boldsymbol{b}_{i})^{2}$\hfill{$\triangleright$ $\boldsymbol{b}_{i}$ is the $i^{\text{th}}$ standard basis}
 \STATE $B \leftarrow B \cup \{g_{i}\}$
 \STATE $V \leftarrow V_{\perp}$ \hfill{$\triangleright$ an orthonormal basis for the subspace of $V$ orthogonal to $\boldsymbol{b}_{i}$}
 \ENDWHILE
 \RETURN minibatch $B$ with size $k$
\end{algorithmic}
\end{algorithm}

@@

Alg.~\ref{alg:goal_sample} shows the details of the goal selection subroutine, and Alg.~\ref{alg:main_alg} gives the overall algorithm for our method, DTGSH.

\section{Experiments}
We evaluate our proposed method, and compare it with current state-of-the-art HER-based algorithms~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy} on challenging robotic manipulation tasks~\cite{plappert2018multi}, pictured in Fig.~\ref{fig:env}. Furthermore, we perform ablation studies on our diversity-based trajectory and goal selection modules. Our code is based on OpenAI Baselines\footnote[1]{\url{https://github.com/openai/baselines}}, and is available at: \url{https://github.com/TianhongDai/div-hindsight}.

\begin{figure}
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/push_resize.png}
    \caption{Push}
    \label{subfig:env_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/pick_resize.png}
    \caption{Pick$\&$Place}
    \label{subfig:env_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handegg_resize.png}
    \caption{EggFull}
    \label{subfig:env_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handblock_resize.png}
    \caption{BlockRotate}
    \label{subfig:env_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handpen_resize.png}
    \caption{PenRotate}
    \label{subfig:env_handpen}
  \end{subfigure}\hfill
  \caption{Robotic manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
  \label{fig:env}
\end{figure}

\subsection{Environments}
The robotic manipulation environments used for training and evaluation include five different tasks. Two tasks use the 7-DoF Fetch robotic arm with two-fingers parallel gripper: Push, and Pick\&Place, which both require the agent to move a cube to the target position. The remaining three tasks use a 24-DoF Shadow Dexterous Hand to manipulate an egg, a block and a pen, respectively. The sparse reward function is given by Eq.~\eqref{eq:sparse_reward}.

In the Fetch environments, the state $s_{t}$ contains the position and velocity of the joints, and the position and rotation of the cube. Each action $a_{t}$ is a 4-dimensional vector, with three dimensions specifying the relative position of the gripper, and the final dimension specifying the state of the gripper (i.e., open or closed). The desired goal $g$ is the target position, and the achieved goal $g^{ac}_{t}$ is the position of the cube. Each episode is of length $T = 50$.

In the Shadow Dexterous Hand environments, the state $s_{t}$ contains the position and velocity of the joints. Each action $a_{t}$ is a 20-dimensional vector which specifies the absolute position of 20 non-coupled joints in the hand. The desired goal $g$ and achieved goal $g^{ac}_t$ specify the rotation of the object for the block and pen tasks, and the position + rotation of the object for the egg task. Each episode is of length $T = 200$.

\subsection{Training Settings}
We base our training setup on CHER~\cite{fang2019curriculum}. We train all agents on minibatches of size $k = 64$ for 50 epochs using MPI for parallelisation over 16 CPU cores; each epoch consists of 1600 ($16 \times 100$) episodes, with evaluation over 160 ($16 \times 10$) episodes at the end of each epoch. Remaining hyperparameters for the baselines are taken from the original work~\cite{NIPS2017_453fadbd,fang2019curriculum,zhao2018energy}. Our method, DTGSH, uses partial trajectories of length $b = 2$ and $m = 100$ as the number of candidate goals.

\subsection{Benchmark Results}
We compare DTGSH to DDPG~\cite{lillicrapContinuous}, DDPG+HER~\cite{NIPS2017_453fadbd}, DDPG+HEBP~\cite{zhao2018energy} and DDPG+CHER~\cite{fang2019curriculum}. Evaluation results are given based on repeated runs with 5 different seeds; we plot the median success rate with upper and lower bounds given by the \nth{75} and \nth{25} percentiles, respectively.

>>>>

\chapter{Efficient Sampling Strategy with Hindsight}
\label{ch:dtgsh}
\section{Introduction}
% Deep reinforcement learning (DRL)~\cite{arulkumaran2017deep}, in which neural networks are used as function approximators for reinforcement learning (RL), has been shown to be capable of solving complex control problems in several environments, including board games~\cite{schrittwieser2020mastering,silver2017mastering}, video games~\cite{berner2019dota,mnih2015human,vinyals2019grandmaster}, simulated and real robotic manipulation~\cite{andrychowicz2020learning,gu2017deep,levine2016end} and simulated autonomous driving~\cite{kiran2021deep}.

% However, learning from a \textit{sparse} reward signal, where the only reward is provided upon the completion of a task, still remains difficult. An agent may rarely or never encounter positive examples from which to learn in a sparse-reward environment. Many domains therefore provide dense reward signals~\cite{brockman2016openai}, or practitioners may turn to reward shaping~\cite{ng1999theory}. Designing dense reward functions typically requires prior domain knowledge, making this approach difficult to generalise across different environments.

% Fortunately, a common scenario is goal-oriented RL, where the RL agent is tasked with solving different goals within the same environment~\cite{kaelbling1993learning,schaul2015universal}. Even if each task has a sparse reward, the agent ideally \textit{generalises} across goals, making the learning process easier. For example, in a robotic manipulation task, the goal during a single episode would be to achieve a specific position of a target object.
In the last chapter, we have successfully combined hindsight goal relabelling with a state-of-the-art on-policy RL algorithm and introduced a new goal-oriented RL method -- episodic self-imitation learning (ESIL). Although the experiments show the effectiveness of the proposed method, ESIL still has worse sample efficiency than the vanilla hindsight experience replay (i.e., DDPG+HER). This is because DDPG is an off-policy RL algorithm, and it employs a replay buffer to store transitions during training, which enables the usage of past experiences. In this chapter, we focus on finding the shortcoming of HER and improving it to elevate learning efficiency and performance.

As a recap, hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is proposed to improve the learning efficiency of goal-oriented RL agents in sparse reward settings: when past experiences are replayed to train the agent, the desired goal is replaced (in ``hindsight'') with the achieved goal, generating many positive experiences. For example, in a robot manipulation task, the goal during a single episode would be to achieve a specific position of a target object. After applying HER, the desired target position would be overwritten with the achieved target position, with the achieved reward also being modified correspondingly.

We note that HER, whilst it enables solutions to previously unsolved tasks, can be somewhat inefficient in its use of uniformly sampling transitions during training. In the same way that prioritised experience replay~\cite{schaul2016prioritized} has significantly improved over the standard experience replay in RL, several approaches have improved upon HER by using data-dependent sampling~\cite{fang2019curriculum,zhao2018energy}. HER with energy-based prioritisation (HEBP)~\cite{zhao2018energy} assumes semantic knowledge about the goal-space and uses the energy of the target objects to sample trajectories with high energies, and then samples transitions uniformly. Curriculum-guided HER (CHER)~\cite{fang2019curriculum} samples trajectories uniformly, and then samples transitions based on a mixture of proximity to the desired goal and the diversity of the samples; CHER adapts the weighting of these factors over time. In this chapter, we introduce diversity-based trajectory and goal selection with HER (DTGSH; See Figure~\ref{fig:illustration_c4}), which samples trajectories based on the diversity of the goals achieved within the trajectory, and then samples transitions based on the diversity of the set of samples. In the experiments, DTGSH is evaluated on five challenging robot manipulation tasks. From extensive experiments, our proposed method converges faster and reaches higher rewards than prior work, without requiring domain knowledge~\cite{zhao2018energy} or tuning a curriculum~\cite{fang2019curriculum}, and is based on a single concept---determinantal point processes (DPPs)~\cite{kulesza2012determinantal}.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/chapter4/illustration_DTGSH_latest.pdf}
    \caption{Overview of DTGSH. Every time a new episode is completed, its diversity is calculated, and it is stored in the episodic replay buffer. During training, $m$ episodes are sampled according to their diversity-based priority, and then $k$ diverse, hindsight-relabelled transitions are sampled using a $k$-DPP~\cite{kulesza2011k}.}
    \label{fig:illustration_c4}
\end{figure}

\section{Related Work}
The proposed work is built on HER~\cite{andrychowicz2017hindsight} as a way to effectively augment goal-oriented transitions from a replay buffer: to address the problem of sparse rewards, transitions from unsuccessful trajectories are turned into successful ones. HER uses an episodic replay buffer, with uniform sampling over trajectories, and uniform sampling over transitions. However, these samples may be redundant, and many may contribute little to the successful training of the agent.

In the literature, some efforts have been made to increase the efficiency of HER by prioritising more valuable episodes/transitions. Motivated by the work-energy principle in physics, HEBP~\cite{zhao2018energy} assigns higher probability to trajectories in which the target object has higher energy; once the episodes are sampled, the transitions are then sampled uniformly. However, HEBP requires knowing the semantics of the goal space in order to calculate the probability, which is proportional to the sum of the target's potential, kinetic and rotational energies.

CHER~\cite{fang2019curriculum} dynamically controls the sampling of transitions during training based on a mixture of goal proximity and diversity. Firstly, $m$ episodes are uniformly sampled from the episodic replay buffer, and then a minibatch of $k < m$ is sampled according to the current state of the curriculum. The curriculum initially biases sampling to achieved goals that are close to the desired goal (requiring a distance function), and later biases sampling towards diverse goals, using a $k$-nearest neighbour graph and a submodular function to more efficiently sample a diverse subset of goals (using the same distance function).

Other work has expanded HER in orthogonal directions. Hindsight policy gradient \cite{rauber2018hindsight} and episodic self-imitation learning~\cite{dai2020episodic} apply HER to improve the efficiency of goal-oriented on-policy algorithms. Dynamic HER~\cite{fang2018dher} and competitive ER~\cite{liu2018competitive} expand HER to the dynamic goal and multi-agent settings, respectively.

The use of DPPs in RL has been more limited, with applications towards modelling value functions of sets of agents in multiagent RL~\cite{osogami2019determinantal,pmlr-v119-yang20i}, and most closely related to us, finding diverse policies~\cite{parker2020effective}.

\section{Methodology}
We now formally describe the two main components of our method, DTGSH: 1) a diversity-based trajectory selection module to sample valuable trajectories for the further goal selection, and 2) a diversity-based goal selection module to select transitions with diverse goal states from the previously selected trajectories. Together, these select informative transitions from a large area of the goal space, improving the agent's ability to learn and generalise.

\subsection{Diversity-based Trajectory Selection}
We propose a diversity-based prioritisation method to select valuable trajectories for efficient training. Related to HEBP's prioritisation of high-energy trajectories~\cite{zhao2018energy}, we hypothesise that trajectories that achieve diverse goal states $g^{ac}_{t}$ are more valuable for training; however, unlike HEBP, we do not require knowledge of the goal space semantics.

In a robot manipulation task, the agent needs to move a target object from its initial position, $g^{ac}_{1}$, to the target position, $g$. If the agent never moves the object, despite hindsight relabelling it will not be learning information that would directly help in task completion. On the other hand, if the object moves a lot, hindsight relabelling will help the agent learn about meaningful interactions.

In our approach, DPPs are used to model the diversity of achieved goal states $g^{ac}_{t}$ in an episode, or subsets thereof. For a single trajectory $\Tau$ of length $T$, we divide it into several partial trajectories $\tau_{j}$ of length $b$, with achieved goal states $\{g^{ac}_{t}\}_{t=n:n+b-1}$. That is, with a sliding window of $b = 2$, a trajectory $\Tau$ can be divided into $N_p$ partial trajectories:
\begin{equation}
    \Tau_{i} = \{\{\underbrace{g^{ac}_{1}, g^{ac}_{2}}_{\tau_{1}}\}, \{\underbrace{g^{ac}_{2}, g^{ac}_{3}}_{\tau_{2}}\}, \{\underbrace{g^{ac}_{3}, g^{ac}_{4}}_{\tau_{3}}\}, \cdots, \{\underbrace{g^{ac}_{T-1}, g^{ac}_{T}}_{\tau_{N_{p}}}\}\}.
\end{equation}
The diversity $d_{\tau_{j}}$ of each partial trajectory $\tau_{j}$ can be computed as:
\begin{equation}
    d_{\tau_{j}} = \text{det}(\mathbf{L}_{\tau_{j}}), 
\label{eq:partial_div}
\end{equation}
where $\mathbf{L}_{\tau_{j}}$ is the kernel matrix of partial trajectory $\tau_{j}$:
\begin{equation}
    L_{\tau_{j}} = \mathbf{M}^{T}\mathbf{M},
\end{equation}
and $\mathbf{M}=[\hat{g}^{ac}_{n}, \hat{g}^{ac}_{n+1}, \cdots, \hat{g}^{ac}_{n+b-1}]$, where each $\hat{g}^{ac}$ is the $\ell_2$-normalised version of the achieved goal $g^{ac}$~\cite{kulesza2011k}. Finally, the diversity $d_\Tau$ of trajectory $\Tau$ is the sum of the diversity of its $N_p$ constituent partial trajectories:
\begin{equation}
    d_\Tau = \sum_{j=1}^{N_{p}} d_{\tau_{j}}.
\label{eq:total_div}
\end{equation}

@@

\subsection{Diversity-based Goal Selection}
In prior work~\cite{andrychowicz2017hindsight,zhao2018energy}, after selecting the trajectories from the replay buffer, one transition from each selected trajectory is sampled uniformly to construct a minibatch for training. However, the modified goals $g^{\prime}$ in the minibatch might be similar, resulting in redundant information. In order to form a minibatch with diverse goals for more efficient learning, we use $k$-DPPs~\cite{kulesza2011k} for sampling goals. Compared to the standard DPP, a $k$-DPP is a conditional DPP where the subset $Y$ has a fixed size $k$, with the probability distribution function:
\begin{equation}
    \mathcal{P}_{L}^{k}(\mathbf{Y}=Y) = \frac{\text{det}(\mathbf{L}_{Y})}{\sum_{|Y^{\prime}|=k} \text{det}(\mathbf{L}_{Y^{\prime}})}.
\end{equation}
$k$-DPPs are more appropriate for goal selection with a minibatch of fixed size $k$. Given $m > k$ trajectories sampled from the replay buffer, we first uniformly sample a transition from each of the $m$ trajectories. Finally, a $k$-DPP is used to sample a diverse set of transitions based on the relabelled goals $g'$ (which, in this context, we denote as ``candidate goals''). Figure~\ref{fig:goal_sampling} gives an example of uniform vs. $k$-DPP sampling, demonstrating the increased coverage of the latter. Figure~\ref{fig:kde} provides corresponding estimated density plots; note that the density of the $k$-DPP samples is actually more uniform over the \textit{support} of the candidate goal distribution.
\begin{figure}[h]
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/goals_distribution_2d.pdf}
        \caption{Plot of candidate goals and selected goals. $k$-DPP sampling is more likely to sample points from the full span of the goal space.}
        \label{fig:goal_sampling}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/chapter4/kde.pdf}
        \caption{Kernel density estimation of the distributions of goals. $k$-DPP leads to a more uniform selection of goals over the support of the goal space.}
        \label{fig:kde}
    \end{subfigure}
    \caption{Visualisation of $k=32$ goals selected from $m=100$ candidate goals of the Push task using either uniform sampling or $k$-DPP sampling, respectively. The candidate goals are distributed over a 2D ($xy$) space. Note that  $k$-DPP sampling (right hand plots) results in a broader span of selected goals in $xy$ space compared to uniform sampling (left hand plots).}
\end{figure}

% pseudo code of dpp-sampling
\begin{algorithm}[t!]
\caption{Diversity-based Goal Selection using $k$-DPP}
\label{alg:goal_sample}
\begin{algorithmic}[1]
 \REQUIRE set of $m$ candidate goal states $\mathcal{G}:=\{g_{i}\}_{i=1:m}$, minibatch size $k$
 \STATE $J \leftarrow\varnothing$, $\mathbf{M} \leftarrow  [g_{1}, g_{2}, \cdots, g_{m}]$
 \STATE Calculate the DPP kernel matrix $\mathbf{L}_\mathbf{M}$
 \STATE $\{\boldsymbol{v}_{n}, \lambda_{n}\} \leftarrow \text{EigenDecomposition}(\mathbf{L}_\mathbf{M})$
 %\STATE $e_{k}(\lambda_{1}, \cdots, \lambda_{m}): = \sum\limits_{|J'|=k}\prod_{n\in J'}\lambda_{n}$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \STATE $e_k(\lambda_1,\lambda_2,\dots,\lambda_m) := 
  \sum\limits_{J^{\prime} \subseteq \{1,2,\dots,m\}\atop|J^{\prime}|=k} \prod\limits_{n \in J^{\prime}} \lambda_n$ \hfill{$\triangleright$ elementary symmetric polynomial: $e^{m}_{k}$}
 \FOR{$\mathrm{n} = m, m-1, \cdots, 1$}
 \IF{$u\sim \text{Uniform}[0, 1] < \lambda_{n}\frac{e^{n-1}_{k-1}}{e_{k}^{n}}$}
 \STATE $J \leftarrow J \cup \{n\}$, $k \leftarrow k - 1$
 \IF{$k = 0$}
 \STATE \textbf{break}
 \ENDIF
 \ENDIF
 \ENDFOR
 \STATE $V \leftarrow \{\boldsymbol{v}_{n}\}_{n \in J}$, $B \leftarrow \varnothing$
 \WHILE{$|V| > 0$}
 \STATE Select $g_{i}$ from $\mathcal{G}$ with $p(g_{i}):=\frac{1}{|V|}\sum_{\boldsymbol{v}\in V}(\boldsymbol{v}^{T}\boldsymbol{b}_{i})^{2}$\hfill{$\triangleright$ $\boldsymbol{b}_{i}$ is the $i^{\text{th}}$ standard basis}
 \STATE $B \leftarrow B \cup \{g_{i}\}$
 \STATE $V \leftarrow V_{\perp}$ \hfill{$\triangleright$ an orthonormal basis for the subspace of $V$ orthogonal to $\boldsymbol{b}_{i}$}
 \ENDWHILE
 \RETURN minibatch $B$ with size $k$
\end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{alg:goal_sample} shows the details of the goal selection subroutine, and Algorithm~\ref{alg:main_alg} gives the overall algorithm for our method, DTGSH.

\section{Experiments}
We evaluate our proposed method, and compare it with current state-of-the-art HER-based algorithms~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy} on challenging robot manipulation tasks~\cite{plappert2018multi}, pictured in Figure~\ref{fig:env_c4}. Furthermore, we perform ablation studies on our diversity-based trajectory and goal selection modules. Our code is based on OpenAI Baselines\footnote[1]{\url{https://github.com/openai/baselines}}, and is available at: \url{https://github.com/TianhongDai/div-hindsight}.

\begin{figure}
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/push_resize.png}
    \caption{Push}
    \label{subfig:env_push}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/pick_resize.png}
    \caption{Pick$\&$Place}
    \label{subfig:env_pick}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handegg_resize.png}
    \caption{EggFull}
    \label{subfig:env_handegg}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handblock_resize.png}
    \caption{BlockRotate}
    \label{subfig:env_handblock}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.19\textwidth}
    \includegraphics[width=\textwidth]{figures/chapter4/handpen_resize.png}
    \caption{PenRotate}
    \label{subfig:env_handpen}
  \end{subfigure}\hfill
  \caption{Robot manipulation environments. (a-b) use the Fetch robot, and (c-e) use the Shadow Dexterous Hand.} 
  \label{fig:env_c4}
\end{figure}

\subsection{Environments}
The robot manipulation environments used for training and evaluation includes five different tasks. Two tasks use the 7-DoF Fetch robot arm with two-fingers parallel gripper: Push, and Pick\&Place, which both require the agent to move a cube to the target position. The remaining three tasks use a 24-DoF Shadow Dexterous Hand to manipulate an egg, a block and a pen, respectively. The sparse reward function is given by Equation~\eqref{eq:sparse_reward}.

In the Fetch environment, the state $s_{t}$ contains the position and velocity of the joints, and the position and rotation of the cube. Each action, $a_{t}$, is a 4-dimensional vector, with three dimensions specifying the relative position of the gripper, and the final dimension specifying the state of the gripper. The desired goal $g$ is the target position, and the achieved goal $g^{ac}_{t}$ is the position of the cube. Each episode is of length $T = 50$.
%(i.e., open or closed)

In the Shadow Dexterous Hand environment, the state $s_{t}$ contains the position and velocity of the joints. Each action $a_{t}$ is a 20-dimensional vector which specifies the absolute position of 20 non-coupled joints in the hand. The desired goal, $g$, and achieved goal, $g^{ac}_t$, specify the rotation of the object for the block and pen tasks, and the position + rotation of the object for the egg task. Each episode is of length $T = 200$.

\subsection{Experimental Settings}
We base our training setup on CHER~\cite{fang2019curriculum}. We train all agents on minibatches of size $k = 64$ for 50 epochs using MPI for parallelisation over 16 CPU cores; each epoch consists of 1600 ($16 \times 100$) episodes, with evaluation over 160 ($16 \times 10$) episodes at the end of each epoch. Remaining hyperparameters for the baselines are taken from the original respective works~\cite{andrychowicz2017hindsight,fang2019curriculum,zhao2018energy}. Our method, DTGSH, uses partial trajectories of length $b = 2$ and $m = 100$ as the number of candidate goals.

\subsection{Results on Benchmark Environments}
We compare DTGSH to DDPG~\cite{lillicrapContinuous}, DDPG+HER~\cite{andrychowicz2017hindsight}, DDPG+HEBP~\cite{zhao2018energy} and DDPG+\\CHER~\cite{fang2019curriculum}. Evaluation results are given based on repeated runs with 5 different seeds; we plot the median success rate with upper and lower bounds given by the \nth{75} and \nth{25} percentiles, respectively.
